Learning Causally-Generated Stationary Time Series

### 2017-02-24 doc/paper.tex
Extra remark

Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Rearranging
\begin{align*}
    &\log p(e) - \KL(q(f,u,z)\|p(f,u,z\cond e)) \\
    &\quad= \ll\la \log \frac{p(e\cond f)\cancel{p(f\cond u, z)}p(u)p(z)}{\cancel{p(f\cond u, z)}q(u)q(z)} \rr\ra_{q(f,u,z)} \\
    &\quad= \underbrace{\la \log p(e\cond f) \ra_{q(f,u,z)}}_{\text{reconstruction cost}} \\
    &\quad\phan{=}- \underbrace{\ll(\ll\la\log\frac{q(u)}{p(u)}\rr\ra_{q(u)} + \ll\la\log\frac{q(z)}{p(z)}\rr\ra_{q(z)}\rr)}_{\text{divergence from prior}} \\
    &\quad=\mathcal{L}[q(u),q(z)]
\end{align*}
shows that we can find $q(u)$ and $q(z)$ through maximising $\mathcal{L}$. Since $\log p(e)\ge\mathcal{L}$, $\mathcal{L}$ is called the \textit{evidence lower bound} (ELBO). Observe that maximisation of $\mathcal{L}$ attempts to explain the data well whilst not diverging too far from the model prior.

@@

\Cref{fig:toy_acausal_sample} shows the results for fitting the GPCM and CGPCM to a noisy random sample from the GPCM. We make several observations. First, the GPCM and CGPCM appear to both succesfully fit the function $f \cond h$ and the kernel $k_{f\cond h}$. This shows that the CGPCM can indeed learn kernels. Second, the sample from the GPCM appears smooth, which means that the learned filter $h$ for the CGPCM should be nearly zero near zero. A quick inspection of \cref{fig:toy_acausal_sample} verifies that this is indeed the case. Finally, the learned filter $h(t)$ for the GPCM appears uncertain for $t\le 0$. This a direct consequence of the causality contraint: \cref{eq:model} shows that $f\cond h$ does not depend on $h(t)$ for $t \le 0$.

@@

Finally, note that the SMF approximation consistently has better calibrated credible regions compared to the MF approximation. For example, in the filter $h$, kernel $k_{f\cond h}$, PSD of $f \cond h$ of the CGPCM in \cref{fig:toy_causal_sample}, the MF approximation fails to capture the truth in its credible region, whereas the SMF approximation does capture the truth.
% \subsection{Speech}
% Coming up.

>>>>

Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Contrary to the formulation of \citet{Tobar:2015:Learning_Stationary}, we have analytically integrated out $h$ and $x$ in $p(f\cond u, z)=\int p(f, h, x\cond u, z)\id{h}\id{x}$ prior to performing inference. It is indeed the case that $p(f\cond u, z)$ is intractable, but its first two moment can actually be computed, and only those we need in formulating our inference scheme. Rearranging
\begin{align*}
    &\log p(e) - \KL(q(f,u,z)\|p(f,u,z\cond e)) \\
    &\quad= \ll\la \log \frac{p(e\cond f)\cancel{p(f\cond u, z)}p(u)p(z)}{\cancel{p(f\cond u, z)}q(u)q(z)} \rr\ra_{q(f,u,z)} \\
    &\quad= \underbrace{\la \log p(e\cond f) \ra_{q(f,u,z)}}_{\text{reconstruction cost}} \\
    &\quad\phan{=}- \underbrace{\ll(\ll\la\log\frac{q(u)}{p(u)}\rr\ra_{q(u)} + \ll\la\log\frac{q(z)}{p(z)}\rr\ra_{q(z)}\rr)}_{\text{divergence from prior}} \\
    &\quad=\mathcal{L}[q(u),q(z)]
\end{align*}
shows that we can find $q(u)$ and $q(z)$ through maximising $\mathcal{L}$. Since $\log p(e)\ge\mathcal{L}$, $\mathcal{L}$ is called the \textit{evidence lower bound} (ELBO). Observe that maximisation of $\mathcal{L}$ attempts to explain the data well whilst not diverging too far from the model prior.

@@

\Cref{fig:toy_acausal_sample} shows the results for fitting the GPCM and CGPCM to a noisy random sample from the GPCM. We make several observations. First, the GPCM and CGPCM appear to both succesfully fit the function $f \cond h$ and the kernel $k_{f\cond h}$. This shows that the CGPCM can indeed learn kernels. Second, the sample from the GPCM appears smooth, which means that the learned filter $h$ for the CGPCM should be nearly zero near zero. A quick inspection of \cref{fig:toy_acausal_sample} verifies that this is indeed the case. Finally, the learned filter $h(t)$ for the CGPCM appears uncertain for $t\le 0$. This a direct consequence of the causality contraint: \cref{eq:model} shows that $f\cond h$ does not depend on $h(t)$ for $t \le 0$.

@@

Finally, note that the SMF approximation consistently has better calibrated credible regions compared to the MF approximation. For example, the MF approximation fails to capture the truth in its credible region for the filter $h$, kernel $k_{f\cond h}$, and PSD of $f \cond h$ of the CGPCM in \cref{fig:toy_causal_sample}; in contrast, the SMF approximation does capture the truth.
% \subsection{Speech}
% Coming up.
